Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-24 TACAS2018/paper.tex
changed CEGORS to CEGIS-OR

We build upon this paradigm to create CEGORS (Counter Example Guided, Optimization
Refined Synthesis) \note{not sure whether we need an acronym for this, we could just call it
CEGIS with optimization and abstraction refinement}.
Our approach is illustrated in Figure~\ref{fig:CEGORS} and consists of three main
phases:
  \note{explain broadly what each phase does}
\begin{itemize}
\item {\sc Abstract}: constructs of an unsound abstract model that over-approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer.  Since
  we are using coarse approximations, these do not necessarily meet any portion
  of the specification, but either through optimization or refinement become
  more precise (therefore more likely to meet the specification) with each iteration.
\item {\sc model}: refines or redefines the abstractions needed by {\sc Abstract}
  in order to reduce the probability of failure in the next iteration.
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon.
\end{itemize}

@@

%Our CEGORS loop is similar to CEGIS with solution generalization~\cite{david2017program}.

@@

The CEGORS loop for this approach is described next.
\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:observable}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
\note{This paragraph needs to be moved before the itemize and it needs to be improved: give an intuition for what $\phi_{static}$ and $\phi_{synth}$ represent, e.g. the system's abstraction + stability spec + safety spec...}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

@@

 \end{tikzpicture}
}
\caption{CEGIS with Optimization Refinement. Orange boxes (light) are modelling
stages, blue boxes (dark) are verification stages, and the red box is the synthesis stage.}
\label{fig:CEGORS}
\end{figure}

@@

\note{remove some of the subcases of CEGIS and CEGORS. Those that are still used need to be explained in the actual section, not in the table's caption.}

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGORS} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max}\\\hline
1  & DC Motor          & 2 & 100 & 9.5\,s & 1.9\,s& \xmark& 1.2\,s &   3.1\,s &   14.1\,s\\
2  & Helicopter        & 3  & 100 & \xmark & \xmark & \xmark & 6.7\,s  &   18.9\,s & \xmark\\
3  & Inverted Pendulum & 4 & 100 &   9.7\,s & \xmark &\xmark & 12.0\,s  &  18.1\,s &  414.0\,s\\
4  & Magnetic Pointer  & 2  & 100 & \xmark & \xmark & \xmark& 4.6\,s  &  39.0\,s &  \xmark \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 14.0\,s & 2.0\,s & \xmark&1.2\,s  &   1.5\,s &   4.3\,s\\
6  & Suspension        & 4 & 100 & 73.7\,s& \xmark & \xmark &24.0\,s  &   46.1\,s &  557.9\,s\\
7  & Tape Driver       & 3 & 100 & 10.1\,s& 140.3\,s & \xmark &2.6\,s  &   10.5\,s &   104.6\,s\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, Maximum and Median times for standard CEGIS vs CEGORS on multiple benchmarks. \xmark~indicates failure to find a controller, either due to timeout or insufficient numeric precision (bitvectors are limited to 64bits).}
\end{table}

@@

The smaller benchmarks (order=2) tend to show little or no improvement when we
apply CEGORS.  The reason is that for these benchmarks the synthesis time using
Solution Generalization is already small, and the overhead of the abstraction is
often equal or larger than the gain obtained.

More interesting cases appear when we increase the order, showing initially that
benchmarks that timed-out using standard CEGIS are now running in under 2
minutes using CEGORS. This is the result of removing several multiplications which are
known to be difficult for SAT solvers.  Furthermore, when we introduce Optimization
Refinement there is a further improvement up to an order of magnitude.  This second
improvement is due to the elimination of counterexamples that would have been
created without the refinement. Since further synthesis require increasingly
complex models which are linear in the number of multiplications, eliminating 
these results in large improvements.  In cases when the initial candidate is
suitable, we gain nothing with the refinement and therefore find similar times.
Improvements using this approach will therefore be observed in higher order
systems or when the safety specification is tight and the original loop would
require multiple counterexamples.

>>>>

We build upon this paradigm to create CEGIS-OR (CEGIS with Optimization
Refinement).
Our approach is illustrated in Figure~\ref{fig:CEGIS-OR} and consists of three main
phases:
  \note{explain broadly what each phase does}
\begin{itemize}
\item {\sc Abstract}: constructs of an unsound abstract model that over-approximates
the dynamics of the system.
\item {\sc synthesize}: generates a candidate controller and observer.  Since
  we are using coarse approximations, these do not necessarily meet any portion
  of the specification, but either through optimization or refinement become
  more precise (therefore more likely to meet the specification) with each iteration.
\item {\sc model}: refines or redefines the abstractions needed by {\sc Abstract}
  in order to reduce the probability of failure in the next iteration.
\item {\sc verify}: verifies the behaviour of the candidate controller and observer
  against a sound over-approximation of the concrete dynamics of the system for an infinite space over an infinite time horizon.
\end{itemize}

@@

%Our CEGIS-OR loop is similar to CEGIS with solution generalization~\cite{david2017program}.

@@

The CEGIS-OR loop for this approach is described next.
\paragraph{1. {\sc abstract}} creates an abstraction from the initial specificationas follows:
\begin{itemize}
\item Select a sample time that minimizes the magnitudes of the matrices used in both the
implementation and the abstraction. This ensures that a small bit vector size is required for
the synthesis process.
\item Build the observer matrices ${\mat{A}_d}_{\langle C \rangle}$, ${\mat{B}_d}_{\langle C \rangle}$
and ${\mat{C}_d}_{\langle C \rangle}$ using the truncation described in Section~\ref{sec:numeric_rep}.
\item Transform the system into canonical form and add nondeterminism for the errors
using the techniques described in Section~\ref{sec:cof_fwl_stability}. The transformation
matrices will be needed to obtain an FWL compliant controller and observer.
\item Compute the characteristic polynomial of the matrix $(\mat{A}_d-\mat{B}_d\mat{K}_{\langle C \rangle})$
as $P_a(z) = z^p+\sum_{i=1}^p{(a_i-k_i)z^{p-i}}$. 
\item Compute the characteristic polynomial of the matrix $({\mat{A}_d}_{\langle C \rangle}-\mat{L}_{\langle C \rangle}{\mat{C}_d}_{\langle C \rangle})$
as $P_l(z) = z^p+\sum_{i=1}^p{(l_i-a_i)z^{p-i}}$. 
\item Create a stability specification based on the characteristic polynomials, perturbation
error ($\mat{A}_e$,$\mat{B}_e$), and response time $\phi_{stability}(P_a,P_l,\overline{\lambda})$
as described in Section~\ref{sec:transientspecification}.
\item Calculate the noise sets $\eta_1$ and $\eta_2$ from the equations
described in Section~\ref{sec:observable}.
\item Calculate a set of initial bounds on $\mat{K}$, $\phi_\mathit{init}^{K}$,
based on the input constraints.
$(\phi_\mathit{init} \wedge \phi_\mathit{input} \wedge u_k=-\mat{K} \vec{x}_k)
\Rightarrow \phi_\mathit{init}^{K}$.
These bounds will be used by the {\sc synthesize} phase to reduce the size of the solution space. 
\item Select an initial width for the bit vectors representing the controller and the plant $\phi_{\langle C \rangle}, \phi_{\langle P \rangle}$. We will increase either precision as needed.
\end{itemize}
\note{This paragraph needs to be moved before the itemize and it needs to be improved: give an intuition for what $\phi_{static}$ and $\phi_{synth}$ represent, e.g. the system's abstraction + stability spec + safety spec...}
There are two abstractions created by these specifications: $\phi_{synth}$ which is given to the synthesiser and $\phi_{static}$ which is given to the synthesis refinement loop. These abstractions will be refined independently, so we will not detail what each will contain at a given stage. We only note that at all times $\phi_{static} \models \phi_{synth}$.
%
%\item
\paragraph{2. {\sc synthesize}} generates a candidate controller $\mat{K}$
and observer $\mat{L}$ that satisfy
$\phi_{synth}$ by invoking a SAT solver.
There are two exit conditions:
If there is no candidate solution we return FAIL and signal {\sc Abstract}. Based on the current abstraction,
{\sc  Abstract} may choose to increase the precision, refine the abstraction, or return a synthesis failure.
If we have a candidate solution we will feed it to the verifier chain in order to validate it.
%
%\item
\paragraph{3. {\sc Verify Synthesis}} verifies $\phi_{static}$ of  both the
controller and the observer.  If one of these results does not satisfy this tighter
specification, it feeds back to the {\sc Optimise} phase.
If the candidate is valid, we feed it to the next {\sc verify} stage, {\sc Discrete Time}.

@@

 \end{tikzpicture}
}
\caption{CEGIS with Optimization Refinement. Orange boxes (light) are modelling
stages, blue boxes (dark) are verification stages, and the red box is the synthesis stage.}
\label{fig:CEGIS-OR}
\end{figure}

@@

\note{remove some of the subcases of CEGIS and CEGIS-OR. Those that are still used need to be explained in the actual section, not in the table's caption.}

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{Number of} & \multicolumn{3}{|c|}{CEGIS} & \multicolumn{3}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{benchmarks}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{Max}\\\hline
1  & DC Motor          & 2 & 100 & 9.5\,s & 1.9\,s& \xmark& 1.2\,s &   3.1\,s &   14.1\,s\\
2  & Helicopter        & 3  & 100 & \xmark & \xmark & \xmark & 6.7\,s  &   18.9\,s & \xmark\\
3  & Inverted Pendulum & 4 & 100 &   9.7\,s & \xmark &\xmark & 12.0\,s  &  18.1\,s &  414.0\,s\\
4  & Magnetic Pointer  & 2  & 100 & \xmark & \xmark & \xmark& 4.6\,s  &  39.0\,s &  \xmark \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 14.0\,s & 2.0\,s & \xmark&1.2\,s  &   1.5\,s &   4.3\,s\\
6  & Suspension        & 4 & 100 & 73.7\,s& \xmark & \xmark &24.0\,s  &   46.1\,s &  557.9\,s\\
7  & Tape Driver       & 3 & 100 & 10.1\,s& 140.3\,s & \xmark &2.6\,s  &   10.5\,s &   104.6\,s\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results. Minimum, Maximum and Median times for standard CEGIS vs CEGIS-OR on multiple benchmarks. \xmark~indicates failure to find a controller, either due to timeout or insufficient numeric precision (bitvectors are limited to 64bits).}
\end{table}

@@

The smaller benchmarks (order=2) tend to show little or no improvement when we
apply CEGIS-OR.  The reason is that for these benchmarks the synthesis time using
Solution Generalization is already small, and the overhead of the abstraction is
often equal or larger than the gain obtained.

More interesting cases appear when we increase the order, showing initially that
benchmarks that timed-out using standard CEGIS are now running in under 2
minutes using CEGIS-OR. This is the result of removing several multiplications which are
known to be difficult for SAT solvers.  Furthermore, when we introduce Optimization
Refinement there is a further improvement up to an order of magnitude.  This second
improvement is due to the elimination of counterexamples that would have been
created without the refinement. Since further synthesis require increasingly
complex models which are linear in the number of multiplications, eliminating 
these results in large improvements.  In cases when the initial candidate is
suitable, we gain nothing with the refinement and therefore find similar times.
Improvements using this approach will therefore be observed in higher order
systems or when the safety specification is tight and the original loop would
require multiple counterexamples.
